Draft-assinment1

### 2017-02-21 Draft-xuezhou-sjb.Rmd
Xiang

#Table of Contents
* 1.[Introduction](#Introduction)
* 2.[Goal](#Goal)
* 3.[Dataset Preparation](#dataprep)
* 4.[Objective](#objective)
* 5.[Summary of variables](#variablesummary)
* 5.1 [Individual variable summary](#5.1)
  * 5.1.1 [School](#5.1.1)
    * 5.1.1.1 [State](#5.1.1.1)
    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
    * 5.1.1.3 [Control of Institution](#5.1.1.3)
  * 5.1.2 Student Body
    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
    * 5.1.2.2 [Number of applications](#5.1.2.2)
  * 5.1.3 Cost 
    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
  * 5.1.4 Financial Aid
    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
  * 5.1.5 Earning
    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
  * 5.1.6 Repayment
    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
* 6.Relationship between variables
 * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
 * 6.2 [The relationship between Debt and Cost](#6.2)
 * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
 * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
 * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
 * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
 * 6.7 [The relationship between PCTLOAN and Control](#6.7)
 * 6.8 [cost and median debts with earning](#6.8)
 * 6.9 [default with earning](#6.9)
 * 6.10 [degree type vs. earning](#6.10)
 * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
 * 6.12 [predominant degree + control (count)](#6.12)
 * 6.13 [control + earning](#6.13)
* 7. [Findings and Conclusion](#7)

@@

The data is collected by United States Department of Education, which corresponds to all undergraduate degree-granting institutions of higher education from 1996 through 2016. The dataset is pretty rich and covers a wide range of topics. It has 1743 associated columns. Each corresponds to one named data variable. All data variables are divided into 10 categories ranging from ‘Root’ category to ‘Earnings’ category. The data type contains integer, string and float. Data elements with an ‘IPEDS’ value in the ‘Source’ field are derived directly from the Integrated Postsecondary Education Data System (IPEDS). For further detailed description of the dataset, you can find it in the CollegeScorecardDataDictionary.xlsx excel file.

#2. Goal<a id="Goal"></a>
College is an investment. It's well-known that a major motivation for students to attend a university is higher post-graduation earning. So will your school make sure you get a return? With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.

@@

state_freq = count(data_clean,"STABBR_factor")[order(-count(data_clean,'STABBR_factor')$freq),]
state_freq
summary(state_freq)
state_freq[1:15,] %>% ggplot(aes(x=STABBR_factor, y=freq)) +
  geom_bar(stat = 'identity')

@@

```{r, message=FALSE}
prop.table(table(data_clean$`PREDDEG_factor`)) * 100
data_clean %>% ggplot(aes(x=PREDDEG_factor, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree") 
```
From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.  43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.

@@

```{r, message=FALSE}
prop.table(table(data_clean$`Control_factor`)) * 100
data_clean %>% ggplot(aes(x=Control_factor, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree")
```

@@

college_app %>%
    ggplot(aes(factor(num_app), perc_student)) +
    geom_boxplot() 
```

@@

earning_clean1[earning_clean1$years_from_entry != 'salary_increase',] %>% 
    ggplot(aes(x = salary, fill =years_from_entry )) +
    geom_density(alpha = 0.5) + 
    xlim(0, 150000)

@@

```{r,message=FALSE}
repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
summary(repayment_clean)
qplot(repayment_clean, geom = "histogram")
```

@@

```{r,message=FALSE}
default_clean = na.omit(data_clean$CDR3)
summary(default_clean)
qplot(default_clean, geom = "histogram")

@@

##6.8 cost and median debts with earning<a id="6.8"></a> 
Next, we analyze the relationship between the amount the debt and post-graduation earning. There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study. We decide debts for students who have not completed to study in this problem, since it makes more sense for the purpose of prediction.

@@

data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN, y = MD_EARN_WNE_P6)) + 
                geom_jitter() +
                geom_smooth(method = "lm", se = FALSE, col = 'red')

@@

```{r, message=FALSE}
data_clean %>% ggplot(aes(x = CDR3, y = MD_EARN_WNE_P6)) + 
                geom_jitter() +
                geom_smooth(method = "lm", se = FALSE, col = 'red')

@@

data_clean %>%
  ggplot(aes(x=`PREDDEG_factor`,
             y=MD_EARN_WNE_P6)) +
  geom_boxplot(aes(fill = `PREDDEG_factor`))

>>>>

#Table of Contents
* 1.[Introduction](#Introduction)

* 2.[Goal](#Goal)

* 3.[Dataset Preparation](#dataprep)
* 4.[Objective](#objective)
* 5.[Summary of variables](#variablesummary)
* 5.1 [Individual variable summary](#5.1)
  * 5.1.1 [School](#5.1.1)
    * 5.1.1.1 [State](#5.1.1.1)
    * 5.1.1.2 [Predominant Degree](#5.1.1.2)
    * 5.1.1.3 [Control of Institution](#5.1.1.3)
  * 5.1.2 Student Body
    * 5.1.2.1 [Percentage first-generation students](#5.1.2.1)
    * 5.1.2.2 [Number of applications](#5.1.2.2)
  * 5.1.3 Cost 
    * 5.1.3.1 [Average net price for Title IV institutions(both public and private institutions)](#5.1.3.1)
    * 5.1.3.2 [Percentage of students who are financially independent](#5.1.3.2)
  * 5.1.4 Financial Aid
    * 5.1.4.1 [Percentage of undergraduates who receive a Pell Grant](#5.1.4.1)
    * 5.1.4.2 [Percent of all federal undergraduate students receiving a federal student loan](#5.1.4.2l)
    * 5.1.4.3 [The median debt for students who have completed](#5.1.4.3)
    * 5.1.4.4 [The median debt for students who have not completed](#5.1.4.4)
  * 5.1.5 Earning
    * 5.1.5.1 [Median family income in real 2015 dollars](#5.1.5.1)
    * 5.1.5.2 [Earning after graduation](#5.1.5.2)
  * 5.1.6 Repayment
    * 5.1.6.1 [3-year Repayment rate](#5.1.6.1)
    * 5.1.6.2 [3-year Default rate](#5.1.6.2)
* 6.Relationship between variables
 * 6.1 [The relationship between NPT4_COMBINE and State](#6.1)
 * 6.2 [The relationship between Debt and Cost](#6.2)
 * 6.3 [The relationship between Percentage of Pell Grant and college type(CONTROL)](#6.3)
 * 6.4 [The relationship between Percentage first-generation students and States](#6.4)
 * 6.5 [The relationship between family income and percentage of Pell Grant](#6.5)
 * 6.6 [The relationship between DEP_STAT_PCT_IND and Control](#6.6)
 * 6.7 [The relationship between PCTLOAN and Control](#6.7)
 * 6.8 [cost and median debts with earning](#6.8)
 * 6.9 [default with earning](#6.9)
 * 6.10 [degree type vs. earning](#6.10)
 * 6.11 [predominant degree, control  and earning after 6 year of entry](#6.11)
 * 6.12 [predominant degree + control (count)](#6.12)
 * 6.13 [control + earning](#6.13)
* 7. [Findings and Conclusion](#7)

@@

These data are collected by United States Department of Education with the goal of helping students and families make more educated decisions about college and compare vital information easily.

#2. Goal<a id="Goal"></a>

College is an investment. It's well-known that a major motivation for students to attend a university is higher post-graduation earning. So will your school make sure you get a return? With collge costs and student-debt levels reaching an all-time high, it's important to know which schools produce the highest earning alumni and what factors can help us predict post-graduate earnings for students in a university.

@@

state_freq = count(data_clean,"STABBR_factor")[order(-count(data_clean,'STABBR_factor')$freq),]
state_freq
summary(state_freq)
state_freq[1:15,] %>% ggplot(aes(x=STABBR_factor, y=freq)) +
  geom_bar(stat = 'identity') +
  xlab('State') +
  ylab('Number of Institution') +
  ggtitle('Number of Institution in Each State') +
  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5))

@@

```{r, message=FALSE}
prop.table(table(data_clean$`PREDDEG_factor`)) * 100
data_clean %>% ggplot(aes(x=PREDDEG_factor, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Degree")  +
  ylab('Count') +
  ggtitle('Number of Institution in Each Predominant Degree type') +
  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5)) 
```
From above frequency table, we see that about 5% of the population are unclassified in terms of the predomiant degree they offer.  43.2% offer Certificate degree as their predominant degree, 27.3% predominantly offer Bachelor degree and only 4.23% are classified as Graduate degree predominant.

@@

```{r, message=FALSE}
prop.table(table(data_clean$`Control_factor`)) * 100
data_clean %>% ggplot(aes(x=Control_factor, fill = 'blue')) + 
  geom_histogram(stat = "count") +
  xlab("Type") +
  ylab('Count') +
  ggtitle('Number of Institution in Each School type') +
  theme(plot.title = element_text(family = "Trebuchet MS", color="#666666", face="bold", size=20, hjust=0.5)) 
```

@@

college_app$num_app = as.factor(college_app$num_app)
levels(college_app$num_app)[levels(college_app$num_app)=="college_app_1"] <- "one"

college_app %>%
    ggplot(aes(factor(num_app), perc_student)) +
    geom_boxplot() +
    xlab("Number of application submitted") +
    ylab("Percentage of Student") +
    labs(title="Percentage of Students Submitting One/Two or Greater Applications") +
    theme(plot.title = element_text(hjust = 0.5))

@@

earning_clean1[earning_clean1$years_from_entry != 'salary_increase',] %>% 
    ggplot(aes(x = salary, fill =years_from_entry )) +
    geom_density(alpha = 0.5) + 
    xlim(0, 150000) +
    xlab("Median Earning") +
    scale_fill_discrete(name="Years from Entry",
                         breaks=c("MD_EARN_WNE_P10", "MD_EARN_WNE_P6"),
                         labels=c("10 years", "6 years"))

@@

```{r,message=FALSE}
repayment_clean = na.omit(data_clean$RPY_3YR_RT_SUPP)
summary(repayment_clean)
qplot(repayment_clean, geom = "histogram") +
xlab("Repayment Rate") +
ylab("Count") +
labs(title="Histogram for Repayment Rate") +
theme(plot.title = element_text(hjust = 0.5))

@@

```{r,message=FALSE}
default_clean = na.omit(data_clean$CDR3)
summary(default_clean)
qplot(default_clean, geom = "histogram") +
xlab("Default Rate") +
ylab("Count") +
labs(title="Histogram for Default Rate") +
theme(plot.title = element_text(hjust = 0.5))

@@

##6.8 Median debts with earning<a id="6.8"></a> 
Next, we analyze the relationship between the amount the debt and post-graduation earning. There are two variables in the dataset to summarize the median amount of debt for student body, one for students who already graduated and one for students who have not completed their study. We decide debts for students who have not completed to study in this problem, since it makes more sense for the purpose of prediction.

@@

data_clean %>% ggplot(aes(x = WDRAW_DEBT_MDN, y = MD_EARN_WNE_P6)) + 
                geom_jitter() +
                geom_smooth(method = "lm", se = FALSE, col = 'red') +
                xlab("Median Amount Debt for Students who have not completed") +
                ylab("Earning  6 Years from Entry")

@@

```{r, message=FALSE}
data_clean %>% ggplot(aes(x = CDR3, y = MD_EARN_WNE_P6)) + 
                geom_jitter() +
                geom_smooth(method = "lm", se = FALSE, col = 'red')+
                xlab("Three-year Cohort Default Rate") +
                ylab("Earning  6 Years from Entry")

@@

data_clean %>%
  ggplot(aes(x=`PREDDEG_factor`,
             y=MD_EARN_WNE_P6)) +
  geom_boxplot(aes(fill = `PREDDEG_factor`)) +
  xlab("Degree Type") +
  ylab("Earning  6 Years from Entry") +
  scale_fill_discrete(name="Degree Type")

S
```
